---
title: "Lab 07 - Conveying the right message through visualisation"
author: "Lily Botha"
date: "02/24/2026"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data}
df <- read_csv("~/Documents/GitHub/lab_07_betterviz/kansas_grouped_rolling_avg.csv")
```
### Exercise 1

The plot is misleading because the axes are on different scales. 

```{r recreate-misleading-plot}
ggplot(df, aes(x = date)) +
  geom_line(aes(
    y = ifelse(mask_mandate == "No Mask", rolling_avg * 2, rolling_avg), 
    color = mask_mandate),
  ) +
  scale_y_continuous(
    name = "Mask",
    sec.axis = sec_axis(~ . / 2, name = "No Mask")
  ) +
  labs(
    title = "Kansas COVID-19 7-Day Rolling Average of Daily Cases/Per 100K Population",
    subtitle = "Mask Counties vs. No-Mask Mandate Counties",
    x = NULL,
    color = NULL
  ) +
  theme_minimal()
```

### Exercise 2

```{r better-plot}
ggplot(df, aes(x = date, y = rolling_avg, color = mask_mandate)) +
  geom_line() +
  labs(
    title = "Kansas COVID-19 Case Rates by Mask Mandate Status",
    subtitle = "7-day rolling average of daily cases per 100K",
    x = NULL,
    y = "Daily cases (per 100K)",
    color = NULL
  ) +
  theme_minimal()
```

### Exercise 3

This plot shows that counties with mask mandates consistently has higher COVID case rates than counties without mask mandates. Rates in mask counties start very high and then gradually decline, while no-mask counties remain relatively stable at a lower level throughout. 

### Exercise 4

This plot shows that COVID case rates differed between counties with and without mask mandates during this short time period, but it doesn't provide strong evidence about whether mask wearing reduces COVID spread because many other factors could explain the trends/differences. This doesn't fully align with my knowledge about how mask wearing reduces transmission. 

### Exercise 5

The accurate visualization showed that counties with mask mandates had consistently higher COVID rates during this time period. Counties with mask mandates decline, but counties with no mask mandates stayed stable. Using a single y-axis is important here so the differences can be judged accurately. Also, labeling the axes as "mask" and "no mask" doesn't make sense. 

### Exercise 6

The opposite message would be that mask and no-mask counties have similar rates, or that mask counties didn't have less COVID cases over time. To create a visualization that reflects that, you could isolate a smaller time period where the trends are closer, like near the end, and adjust the y-axis scale to make the differences look even smaller. 

### Exercise 7

```{r opposite-plot}
ggplot(df, aes(x = date, y = rolling_avg, color = mask_mandate)) +
  geom_line() +
  scale_y_continuous(limits = c(0, 40)) + 
  labs(
    title = "Kansas COVID-19 Case Rates by Mask Mandate Status",
    subtitle = "7-day rolling average of daily cases per 100K",
    x = NULL,
    y = "Daily cases (per 100K)",
    color = NULL
  ) +
  theme_minimal()
```
